\textbf{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

### 2023-11-30 manuscript_BiSC_outdoor_AP.tex

Air pollution measurements were conducted during two separate monitoring campaigns. The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021. The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022. During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and hybrid models.

@@

% --- Distribution of the monitoring campaigns --- % 
\subsection{Distribution of monitoring campaigns}

Distribution of the concentration for the BiSCAPE and BiSC-Home campaign for all the pollutants were shown in \hyperref[table1]{Table 1}. For all the air pollutants assessed, concentration varied substantially between pollutants.

%%% Table 1. Outdoor air pollution concentration from BiSC-home and BiSCAPE campaigns (2018-2021)

%%% --- Table 1. Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns --- %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns}
\begin{tabular}{lllllllll}
\toprule
Campaign & Pollutant & Year & Units & mean & sd & p25 & p50 & p75 \\ 
\midrule
BiSC-home & NO\textsubscript{2} & 2018-2021 & \(\mu\)g/m\textsuperscript{3} & 38.8 & 12.4 & 30.0 & 37.5 & 45.8 \\ 
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{25} & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 13.4 & 6.14 & 9.06 & 11.3 & 16.2 \\ 
& BC & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 1.27 & 0.50 & 0.88 & 1.22 & 1.52 \\ 
& Fe & 2021-2022 & \(\mu\)g/m\textsuperscript{3} & 0.25 & 0.13 & 0.15 & 0.21 & 0.32 \\ 
& Cu & 2021-2022 & ng/m\textsuperscript{3} & 7.10 & 3.82  & 4.37 & 6.36 & 8.98 \\ 
& Mn & 2021-2022 & ng/m\textsuperscript{3} & 40.7 & 36.5 & 18.9 & 28.2 & 46.7 \\ 
\bottomrule
\end{tabular}
\end{table}

The spatial distribution of the campaigns were showed in \cref{fig1}.

Since models were built based on different approaches there where not directly comparable due to differences on a temporal basis

Since different approaches were used to built outdoor air pollution models these were not directly comparable due to differences on temporal

The performance of the models varied substantially by pollutant. In table x. performance metrics for the HM are shown. Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE. The performance of the PM25

Performance metrics for LUR, DM and HM are shown in the supplementary materials files.

\
% --- Exposure concentrations --- %
Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in FIGURE X. The exposure range was larger for the LUR model for all the pollutants assessed. Dispersion model predicted lower exposure concentration values than LUR and Hybrid. In addition to that HM show less variability in the exposure concentration.

%%% --- Figure x. Boxplot models comparison --- %%%
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/boxplot_models.png}
\caption{Distribution exposure concentration for the entire pregnancy period \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Correlation between measures and model estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 (figure 3.A).

Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 (figure 3.B).

@@

Correlation between the predicted air pollutants for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three different approaches (LUR, DM and HM). All models show moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}.

>>>>

Air pollution measurements were conducted during two separate monitoring campaigns. The first, named the "BiSC-home" campaign, took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} levels for one week during the first, and third trimester of pregnancy between 2018-2021. The second campaign was carried out in accordance with the ESCAPE guidelines, named "BiSCAPE", and included four distinct measurement periods: three in the winter, summer, and autumn of 2021, and a final one in the winter of 2022. During these periods, levels of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC were measured at 34 locations across Barcelona Metropolitan Area (Barcelona, Esplugues de Llobregat, and L'Hospitalet).These monitoring campaigns were used in the development of the LUR and HM.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Distribution of the monitoring campaigns --- % 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Distribution of monitoring campaigns}

Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants are shown in \hyperref[table1]{Table 1}. NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 \(\mu \text{g/m}^3\) and 97.3 \(\mu \text{g/m}^3\) with a mean (sd) of  38.8 (12.4) \( \mu \text{g/m}^3\). PM\textsubscript{2.5} observed levels ranged between 6.14 \(\mu \text{g/m}^3\) and 35.8 \( \mu \text{g/m}^3\) with a mean (sd) of 13.4 (6.14) \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign ranged between 0.40 \(\mu \text{g/m}^3\) and 2.95 \(  \mu \text{g/m}^3\) with a mean of 1.27 (0.50) \( \mu \text{g/m}^3\). For PM\textsubscript{2.5} constituents concentration values observed during the BiSCAPE campaign varied substantially between elements with a mean of 0.25 for  FE,  7.10 for Cu, and a mean of 40.7 for Mn.

\\
The spatial distribution of the outdoor air pollution measurements campaigns were showed in \cref{fig1}.  Panel Fig 1.A show a high density coverage

%%% ---------------------------------------------------------------- %%%
%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
%%% ---------------------------------------------------------------  %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
\toprule
Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
\midrule
BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\ 
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{25} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
& Fe (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
& Cu (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
& Mn (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
\bottomrule
\end{tabular}
\end{table}

Since different approaches were used to built outdoor air pollution models these were not directly comparable due to differences on temporal basis.

The performance of the models varied substantially by pollutant. In table x. performance metrics for the HM are shown. Hybrid BC model had the best performance of the models assessed, with a higher R\textsuperscript{2} and lower RMSE. Performance metrics for LUR, DM and HM are shown in the supplementary materials.

% --- Exposure concentrations --- %
Predictions from outdoor air pollution concentration exposure during the entire pregnancy for each pollutant and model is shown in FIGURE X. The exposure range was larger for the LUR model for all the pollutants assessed. Dispersion model predicted lower exposure concentration values than LUR and Hybrid. In addition to that HM show less variability in the exposure concentration.

%%% -------------------------------------------%%%%
%%% --- Figure 2. Boxplot models comparison --- %%%
%%% -------------------------------------------%%%%
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/boxplot_all_models_estimates.png}
\caption{Distribution of exposure to air pollution during the entire pregnancy period for each exposure assessment model (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu}, \textit{Zn})}
\end{figure}

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Correlation between measures and model estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 4.A)}. Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 4.B)}. Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure. Nitrogen dioxide

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Comparison/agreement between air pollution estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% --- Figure. Comparison/agreement between moodel predictions estimates --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/NO2_HEX.png}
\end{figure}

% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/PM25_HEX.png}
\end{figure}

% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/BC_HEX.png}
\caption{Comparison/agreement between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
\end{figure}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three different approaches (LUR, DM, and HM). Models show moderate to good performance

spatiotemporal models using three different approaches (LUR, DM and HM). All models show moderate to good performance evaluated by the defined metrics in the \textbf{methods section} and displayed in more detail in the \textbf{Supplementary materials}.
